# multiStockassessment

The `multiStockassessment` package extends the [`stockassessment`](https://github.com/fishfollower/SAM) to allow linked stocks (See e.g. Albertsen, C. M., Nielsen, A., and Thygesen, U. H. (2018) Connecting single-stock assessment models through correlated survival. *ICES Journal of Marine Science*, 75(1), 235-244. doi: [10.1093/icesjms/fsx114](https://dx.doi.org/10.1093/icesjms/fsx114)). The package can be installed with `devtools`:

```r
devtools::install_github("calbertsen/multi_SAM", subdir = "multiStockassessment")
```

## Short example

To use the `multiStockassessment` package, the individual stocks must first be fitted with the `stockassessment` package and combined to a `samset` using the `c(...)` function.

### Preparing data using stockassessment

The code below loads the North Sea cod data from the stockassessment package, fits it, simulates two new data sets and fit those individually.

```{r,results='hide',warning=FALSE}

library(stockassessment)

data(nscodData)
data(nscodConf)
data(nscodParameters)

set.seed(9876)
fit <- sam.fit(nscodData, nscodConf, nscodParameters,sim.condRE = FALSE)

datSim <- simulate(fit,nsim=2)
fitSim <- do.call("c",lapply(datSim,function(x)sam.fit(x,nscodConf, nscodParameters)))

```

### Creating correlation structure

```{r}
library(multiStockassessment)
```

After the simulated data have been fitted and combined by the `c` function, the `multiStockassessment` package can be used to fit a model with correlated survival.

The `suggestCorStructure` function can be used to set up the correlation matrix. The `nAgeClose` argument can be used to construct the banded structures used in the paper. The result is a logical symmetric matrix indicating whether a correlation should be fixed at zero.
The dimension of the matrix is the number of stocks times the number of ages in the data.

For instance, `suggestCorStructure(fitSim,nAgeClose=2)` creates a matrix where ages less than 2 appart are correlated between the stocks. Likewise, 

```{r}
cs <- suggestCorStructure(fitSim,nAgeClose=2)
cs
```

The `suggestCorStructure` function has several options to configure the correlation structure, and the result can be modified by hand for complete freedom.

### Fitting the model

The model is fitted by the `multisam.fit` function. The function requires a `samset` and a correlaiton structure. By default, the correlation structure given models the partial correlations between cohorts. This can be changed to regular correlations by setting `usePartialCors = FALSE`.

```{r,results="hide",warning=FALSE}
obj <- multisam.fit(fitSim,cs)
```

```{r}
obj
```


### Investigating the result

Most plotting and table functions from the `stockassessment` package have a corresponding version in `multiStockassessment`.

The fitted correlation structure can be plotted by

```{r}

corplot(obj)

```

Other plotting functions implemented from the `stockassessment` package are

```{r}
fbarplot(obj)
```

```{r}
ssbplot(obj)
```

```{r}
ssbplot(obj)
```

```{r}
tsbplot(obj)
```

```{r}
recplot(obj)
```

```{r}
catchplot(obj)
```

```{r}
srplot(obj)
```

```{r}
fitplot(obj,1)
```

Note that the `fitplot` function above requires the stock to be specified.

```{r}
fitplot(obj,2)
```


### Comparing models

Models can be compared with the `modeltable` function. For instance, the model above can be compared to individual assessments by

```{r,results="hide",warning=FALSE}

cs2 <- suggestCorStructure(fitSim,nAgeClose=0)
obj2 <- multisam.fit(fitSim,cs2)

```

```{r}
modeltable(c(obj,obj2))
```
